## LOADING DATA INTO DATAFRAMES

```{r}
library(dplyr)

path_data_folder<-"c:/tmp"
airports <- read.csv(paste(path_data_folder , "/data/airports.csv",sep=""))

airport_frequencies <- read.csv(paste(path_data_folder , "/data/airport-frequencies.csv",sep="")) 

runways <- read.csv(paste(path_data_folder , "/data/runways.csv",sep="")) 

```

## 1) SELECT...WHERE

#### 1.1) Show all records
```{r}
airports
```

#### 1.2) Show only three records

```{r}
head(airports,3)
```

#### 1.3) Show id filtered by ident='KLAX'
```{r}
airports %>% filter(ident == "KLAX")%>%select(id)
```

#### 1.4) Show distinct types 
```{r}
airports%>%distinct(type)
```

## 2) SELECT WITH MULTIPLE CONDITIONS

#### 2.1) Multiple condition, first example
```{r}
airports%>%filter(iso_region==c('US-CA')& type=='seaplane_base' )
```
#### 2.2) Multiple condition, second example
```{r}
airports%>%filter(iso_region==c('US-CA')& type=='seaplane_base' )%>%select(c('ident', 'name', 'municipality'))
```

## 3)ORDER BY

#### 3.1) Order by asc

```{r}
airport_frequencies%>%filter(airport_ident=='KLAX')%>%arrange(type)
```

#### 3.2) Order by desc

```{r}
airport_frequencies%>%filter(airport_ident=='KLAX')%>%arrange(desc(type))
```

## 4) IN...NOT IN

#### 4.1) IN several values
```{r}
airports%>%filter(type==c('heliport', 'balloonport'))
```

####4.2) NOT IN several values

```{r}
airports%>%filter(type!=c('heliport', 'balloonport'))
```

## 5) GROUP BY

####5.1) Group by with count and order by ISO_COUNTRY,TYPE

```{r}
airports%>%group_by(iso_country,type)%>% summarise(number = n())
```

####5.2) Group by with count and order by ISO_COUNTRY,COUNT

```{r}
airports%>%group_by(iso_country,type)%>% summarise(number = n())%>%arrange(iso_country,desc(number))
```

# 6) LIMIT

####6.1) Top 10 records
```{r}
airports%>%group_by(iso_country)%>% summarise(number = n())%>%arrange(desc(number))%>%head(n=10)
```
####6.2) Records with an offset of 10
```{r}
airports%>%group_by(iso_country)%>% summarise(number = n())%>%arrange(desc(number))%>%head(20)%>%tail(10) 

```

# 7) HAVING

#### 7.1) Types with count over 1000
```{r}
airports%>%filter(iso_country=='US')%>%group_by(type)%>% summarise(number = n())%>%arrange(desc(number))%>%filter(number>1000)
```

####8.2) Min, masx, mean and median
```{r}
runways%>%summarise(max = max(length_ft,na.rm=TRUE),min=min(length_ft,na.rm=TRUE),mean=mean(length_ft,na.rm=TRUE),median=median(length_ft,na.rm=TRUE))
```


#9) JOIN (INNER, LEFT, RIGHT AND OUTER)

####9.1)INNER JOIN BY AIRPORT REF

```{r}
airport_frequencies%>%inner_join(airports,by=c("airport_ref"="id"))%>%select(airport_ident,type.x,description,frequency_mhz)
```

#10) UNION AND UNION ALL

####10.1) UNION ALL

```{r}
airports_df1<-airports%>%filter(ident==c('KLGB','KLAX'))

```

```{r}
airports_df2<-airports%>%filter(ident==c('KLGB'))
```

```{r}
airports_union_all<-bind_rows(airports_df1, airports_df2)
```


####10.2) UNION
```{r}
airports_union<-airports_union_all[!duplicated(airports_union_all),]
```


#11) EXCEPT

####11.1) EXCEPT BETWEEN TWO DATAFRAMES
```{r}
airports_except<-airports_df1[!(airports_df1$id %in% airports_df2$id ), ]
```

#12) OVER (PARTITION BY)

####12.1)MAX LATITUDE VALUE BY TYPE OF AIRPORT

```{r}
airports%>%group_by(type)%>%summarise(max = max(latitude_deg,na.rm=TRUE))
```

